\bf An observational study of papers published in \emph{PLOS ONE} and studies posted to a trial registry.

### 2021-05-13 manuscript/asa_template/asa_template.Rmd
update manuscript .tex file

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggpubr)
library(knitr)
library(flextable)
library(tidytext)

@@

Two statisticians on this paper (AB and NW) have heard researchers admit that they have copied-and-pasted their statistical methods sections from other papers, regardless of whether they are appropriate. 
The aim of this paper is to use text-mining methods to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
Boilerplate text is that "which can be reused in new contexts or applications without significant changes to the original" [@Wikipedia].
Use of these methods sections indicates that little thought has gone into the statistical analysis.

@@

_PLOS ONE_ is a large open access journal that publishes original research across a wide range of scientific fields. Articles must be in English. Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated areas of expertise. Submissions do not undergo formal statistical review. Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail” [@PLOS]. All reviewers are asked the question: "Has the statistical analysis been performed appropriately and rigorously?", with the possible responses of "Yes", "No" and "I don't know".

@@

_Step 1_: Targeted API searches. API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted the presence of analysis-related terms anywhere in the article. Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. Search terms were intended to be broad whilst keeping search results to a manageable number for full-text review (see Step 2). By allowing terms to appear anywhere in the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume and subject classification(s).

_Step 2_: Partial matching on section headings. Full text XML data for all search results were downloaded and combined into a single dataset, organised by DOI and subsection heading(s). Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’. For records that did not pass this second stage filter, we selected a random sample of XXX records and reviewed where initial search terms appeared in the full-text, to estimate the likely proportion of statistical methods sections that were missed.

@@

We applied the same text cleaning steps to both data sources using functions from the `textclean package' [@textclean]. 
Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting. _R_ code used for data extraction and cleaning is available from https://github.com/agbarnett/stats_section.

@@

Results from the clustering algorithm were visualised by word clouds, which summarised frequently occurring terms associated with topic membership. For both _PLOS ONE_ and ANZCTR datasets, we reviewed the strongest matches to each cluster based on topic coherence.

Within the _PLOS ONE_ dataset, we used the strongest matches to investigate evidence of boilerplate text, by comparing text with other studies assigned to the same topic. Comparisons were carried out by calculating Jaccard similarity scores between studies at the sentence level. We chose the Jaccard score as an intuitive measure, which summarises the similarity between two sentences by the number of words common to each sentence (intersection), divided by the number of words that appeared in either the strongest match or the sentence being compared. Boilerplate text was defined based on word count (plus or minus 3 words) and a Jaccard score of 0.9 or higher.

We also compared similarities in text between pairs of records assigned to each cluster. Similarities were determined from calculating cosine similarity scores, with comparisons limited to the top 500 records per cluster. Higher similarity score represent greater overlap in text content between statistical methods sections. Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.

@@

\begin{landscape}
\begin{table}[]
\centering
\caption{Example boilerplate text from PLOS ONE dataset}
\label{tab:plos-example-boilerplate}
\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
\hline
\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\
& Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\
\hline
3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\
 & All statistical analysis was performed using Graphpad Prism software.\\
 \hline
4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\
& Data are expressed as the mean ± SEM. Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test. P values less than 0.05 were considered significant.\\
\hline
5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\
  & The results are presented as the mean ± SEM. Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA). The difference was considered statistically significant at P$<$0.05.\\
  \hline
6 & All results are expressed as means ± standard deviation (± SD).\\
& Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\
\hline
9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\
 & Data are presented as the mean ± SD. Statistical analysis was performed using the Students unpaired t-test. Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\
 \hline
\end{tabular}
\end{table}
\end{landscape}

@@

Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}. Detailed findings based on cosine similarity scores are provided in Supplementary File 2.

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggpubr)
library(knitr)
library(flextable)
library(tidytext)
library(openxlsx)

@@

Two statisticians on this paper (AB and NW) have heard researchers admit that they have copied-and-pasted their statistical methods sections from other papers, regardless of whether they are appropriate. 
The aim of this paper is to use text-mining methods to analyse the content of statistical methods sections included in common research outputs. Results are then evaluated to estimate the extent that researchers are using cut-and-paste or 'boilerplate' statistical methods sections.
Boilerplate text is that "which can be reused in new contexts or applications without significant changes to the original" [@Wikipedia].
Use of these methods sections indicates that little thought has gone into the statistical analysis.

@@

_PLOS ONE_ is a large open access journal that publishes original research across a wide range of scientific fields. Article submissions are handled by an academic editor who selects peer reviewers based on their self-nominated areas of expertise. Submissions do not undergo formal statistical review. Instead, reviewers are required to assess submissions against several publication criteria, including whether: “Experiments, statistics, and other analyses are performed to a high technical standard and are described in sufficient detail” [@PLOS]. All reviewers are asked the question: "Has the statistical analysis been performed appropriately and rigorously?", with the possible responses of "Yes", "No" and "I don't know".

@@

_Step 1_: Targeted API searches. API searches were completed using the R package ‘rplos’ [@rplos]. Search queries targeted the presence of analysis-related terms anywhere in the article. Search terms combined the words “data” or “statistical” with one of: “analysis”, “analyses”, “method”, “methodology” or “model(l)ing”. By allowing terms to appear anywhere in the article, we accounted for the possibility of relevant text being placed in different sections, for example, in the _Material and Methods_ section versus _Results_. Search results were indexed by a unique Digital Object Identifier (DOI). Attribute data collected per DOI included journal volume and subject classification(s).

_Step 2_: Partial matching on section headings. Full text XML data were downloaded and combined into a single dataset, organised by DOI and subsection heading(s). Since _PLOS ONE_ does not prescribe standardised headings to preface statistical methods sections, we performed partial matching on available headings against frequently used terms in initial search results: ‘Statistical analysis’, ‘Statistical analyses’, ‘Statistical method’, ‘Statistics’, ‘Data analysis’ and ‘Data analyses’.

<!-- For records that did not pass this second stage filter, we selected a random sample of XXX records and reviewed where initial search terms appeared in the full-text, to estimate the likely proportion of statistical methods sections that were missed. -->

@@

Both datasets were cleaned using functions from the `textclean package' [@textclean]. 
Text cleaning aimed to standardise notation and statistical terminology, whilst minimising changes to article style and formatting. _R_ code used for data extraction and cleaning is available from https://github.com/agbarnett/stats_section.

@@

<!-- Clustering results were visualised by word clouds that summarised frequently occurring terms associated with membership to different topics.  -->

Within the _PLOS ONE_ dataset, statistical methods sections with high topic coherence scores were used to investigate evidence of boilerplate text, compared with other studies assigned to the same topic. Comparisons were carried out by calculating Jaccard similarity scores between studies at the sentence level. We chose the Jaccard score as an intuitive measure, which summarises the similarity between two sentences by the number of words common to each sentence (intersection), divided by the number of words that appeared in either the strongest match or the sentence being compared. Boilerplate text was defined based on word count (plus or minus 3 words) and a Jaccard score of 0.9 or higher.

Results were transformed to lower case for the clustering, but examples are given using the original capitalisation.

@@

<!-- \begin{landscape} -->
<!-- \begin{table}[] -->
<!-- \centering -->
<!-- \caption{Example boilerplate text from PLOS ONE dataset} -->
<!-- \label{tab:plos-example-boilerplate} -->
<!-- \begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}} -->
<!-- \hline -->
<!-- \textbf{Topic} & \textbf{Statistical methods section} \\ \hline -->
<!-- 1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\ -->
<!-- & Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\ -->
<!-- \hline -->
<!-- 3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\ -->
<!--  & All statistical analysis was performed using Graphpad Prism software.\\ -->
<!--  \hline -->
<!-- 4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\ -->
<!-- & Data are expressed as the mean ± SEM. Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test. P values less than 0.05 were considered significant.\\ -->
<!-- \hline -->
<!-- 5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\ -->
<!--   & The results are presented as the mean ± SEM. Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA). The difference was considered statistically significant at P$<$0.05.\\ -->
<!--   \hline -->
<!-- 6 & All results are expressed as means ± standard deviation (± SD).\\ -->
<!-- & Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\ -->
<!-- \hline -->
<!-- 9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\ -->
<!--  & Data are presented as the mean ± SD. Statistical analysis was performed using the Students unpaired t-test. Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\ -->
<!--  \hline -->
<!-- \end{tabular} -->
<!-- \end{table} -->
<!-- \end{landscape} -->

```{r}
jaccard.matches = read.xlsx('../../results/plos.jaccard_matches.xlsx')
ftab = jaccard.matches %>% mutate(Topic = paste('Topic',topic)) %>%
  select(Topic,sentences,matches) %>%  arrange(Topic,-matches) %>%
  rename('Statistical methods text' = sentences,
         'Number of matching studies' = matches)
knitr::kable(ftab,format = "latex", caption='Example boilerplate text from PLOS ONE dataset')
```

@@

Examples of boilerplate text for selected topics are presented in Table \ref{tab:plos-example-boilerplate}.

@@

Finally, targeted searches for the term 'statistician' revealed instances across all topics where no methods were specified. Instead, all analyses were the responsibility of an appointed statistician, for example, "A statistician employed by hospital was used" and ``Pilot study at this point will use a statistician professionally to determine sample size calculations as required''.

### 2021-05-13 manuscript/asa_template/asa_template.tex
update manuscript .tex file

Two statisticians on this paper (AB and NW) have heard researchers admit
that they have copied-and-pasted their statistical methods sections from
other papers, regardless of whether they are appropriate. The aim of
this paper is to use text-mining methods to estimate the extent that
researchers are using cut-and-paste or `boilerplate' statistical methods
sections. Boilerplate text is that ``which can be reused in new contexts
or applications without significant changes to the original''
\citep{Wikipedia}. Use of these methods sections indicates that little
thought has gone into the statistical analysis.

@@

We used two openly available data sources to find statistical methods
sections: research articles published in \emph{PLOS ONE} and study
protocols registered on the Australian and New Zealand Clinical Trials
Registry (ANZCTR). Data sources were chosen as examples of common
research outputs that include descriptions of statistical methods that
were used, or are planned, for analysing study outcomes.

@@

\emph{PLOS ONE} is a large open access journal that publishes original
research across a wide range of scientific fields. Articles must be in
English. Article submissions are handled by an academic editor who
selects peer reviewers based on their self-nominated areas of expertise.
Submissions do not undergo formal statistical review. Instead, reviewers
are required to assess submissions against several publication criteria,
including whether: ``Experiments, statistics, and other analyses are
performed to a high technical standard and are described in sufficient
detail'' \citep{PLOS}. All reviewers are asked the question: ``Has the
statistical analysis been performed appropriately and rigorously?'',
with the possible responses of ``Yes'', ``No'' and ``I don't know''.

@@

\emph{Step 1}: Targeted API searches. API searches were completed using
the R package `rplos' \citep{rplos}. Search queries targeted the
presence of analysis-related terms anywhere in the article. Search terms
combined the words ``data'' or ``statistical'' with one of:
``analysis'', ``analyses'', ``method'', ``methodology'' or
``model(l)ing''. Search terms were intended to be broad whilst keeping
search results to a manageable number for full-text review (see Step 2).
By allowing terms to appear anywhere in the article, we accounted for
the possibility of relevant text being placed in different sections, for
example, in the \emph{Material and Methods} section versus
\emph{Results}. Search results were indexed by a unique Digital Object
Identifier (DOI). Attribute data collected per DOI included journal
volume and subject classification(s).

\emph{Step 2}: Partial matching on section headings. Full text XML data
for all search results were downloaded and combined into a single
dataset, organised by DOI and subsection heading(s). Since \emph{PLOS
ONE} does not prescribe standardised headings to preface statistical
methods sections, we performed partial matching on available headings
against frequently used terms in initial search results: `Statistical
analysis', `Statistical analyses', `Statistical method', `Statistics',
`Data analysis' and `Data analyses'. For records that did not pass this
second stage filter, we selected a random sample of XXX records and
reviewed where initial search terms appeared in the full-text, to
estimate the likely proportion of statistical methods sections that were
missed.

@@

We applied the same text cleaning steps to both data sources using
functions from the `textclean package' \citep{textclean}. Text cleaning
aimed to standardise notation and statistical terminology, whilst
minimising changes to article style and formatting. \emph{R} code used
for data extraction and cleaning is available from
\url{https://github.com/agbarnett/stats_section}.

@@

We compiled an extensive list of statistical terms to standardise
descriptions of statistical methods reported across both datasets. An
initial list was compiled by calculating individual word frequencies and
identifying relevant terms that appeared at least 100 times. Further
terms were sourced from index searches of three statistics textbooks
\citep[\citet{Diggle2013},\citet{Bland2015}]{Dobson2018}. The final list
is provided in Supplementary File 1. Plurals (e.g., `chi-squares')
unhyphenated (e.g., `chi square') and combined (e.g.~`chisquare') terms
were transformed to singular, hyphenated form (e.g., `chi-square').
Common statistical tests were also hyphenated (e.g., `hosmer lemeshow'
to `hosmer-lemeshow').

@@

Text clustering is known to face the curse of dimensionality due to the
high number of terms in doc\(\times\)term matrix representation
\citep[\citet{sutanto2018fine}]{aggarwal2012mining}. Therefore
text-based methods based on distance, density or probability face
difficulties
\citep[\citet{mohotti2018corpus},\citet{mohotti2019concept}]{park2018examining}.
Specifically, the distance difference between near and far points
becomes negligible in high-dimensional data \citep{aggarwal2012mining}.
This directly affects the distance-based methods such as
\textit{k}-means \cite{jain} in accurately identifying the common
subgroups. In addition, the sparseness of this high dimensional matrix
representation does not allow to differentiate the user groups based on
density differences
\citep[\citet{mohotti2018efficient}]{mohotti2018corpus}.

Non-negative Matrix factorization (NMF) which maps the high-dimensional
data to a lower-dimensional space has been found to provide an effective
solution by allowing to form clusters in the lower-dimensional space
\citep[\citet{luong2019clustering}]{kim2014algorithms}. NMF approximates
topic vectors in linear manner considering the context of terms.

@@

Results from the clustering algorithm were visualised by word clouds,
which summarised frequently occurring terms associated with topic
membership. For both \emph{PLOS ONE} and ANZCTR datasets, we reviewed
the strongest matches to each cluster based on topic coherence.

Within the \emph{PLOS ONE} dataset, we used the strongest matches to
investigate evidence of boilerplate text, by comparing text with other
studies assigned to the same topic. Comparisons were carried out by
calculating Jaccard similarity scores between studies at the sentence
level. We chose the Jaccard score as an intuitive measure, which
summarises the similarity between two sentences by the number of words
common to each sentence (intersection), divided by the number of words
that appeared in either the strongest match or the sentence being
compared. Boilerplate text was defined based on word count (plus or
minus 3 words) and a Jaccard score of 0.9 or higher.

We also compared similarities in text between pairs of records assigned
to each cluster. Similarities were determined from calculating cosine
similarity scores, with comparisons limited to the top 500 records per
cluster. Higher similarity score represent greater overlap in text
content between statistical methods sections. Results were transformed
to lower case for the clustering, but examples are given using the
original capitalisation.

@@

{\centering \includegraphics[width=0.6\linewidth]{figures/plos.summary}

@@

\caption{\label{fig:plos-n}PLOS summary}\label{fig:unnamed-chunk-3}
\end{figure}

@@

{\centering \includegraphics[width=0.8\linewidth]{figures/plos.wordclouds}

@@

\begin{landscape}
\begin{table}[]
\centering
\caption{Example boilerplate text from PLOS ONE dataset}
\label{tab:plos-example-boilerplate}
\begin{tabular}{p{0.1\linewidth} p{0.9\linewidth}}
\hline
\textbf{Topic} & \textbf{Statistical methods section} \\ \hline
1 & Student’s t-test was used for statistical analysis. A p value of $<$0.05 was considered statistically significant. \\
& Statistical analysis was performed using the Students t-test. P values of less than 0.05 were considered significant.\\
\hline
3 & GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses.\\
 & All statistical analysis was performed using Graphpad Prism software.\\
 \hline
4 & Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons.\\
& Data are expressed as the mean ± SEM. Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test. P values less than 0.05 were considered significant.\\
\hline
5 & SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis.\\
  & The results are presented as the mean ± SEM. Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA). The difference was considered statistically significant at P$<$0.05.\\
  \hline
6 & All results are expressed as means ± standard deviation (± SD).\\
& Data are expressed as mean ± Standard error of the mean (SEM). Statistical analysis of the data was performed by the Student’s t-test. A value of P$<$0.05 was considered statistically significant.\\
\hline
9 & Statistical significance was determined by Students t-tests. $^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$.\\
 & Data are presented as the mean ± SD. Statistical analysis was performed using the Students unpaired t-test. Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$.\\
 \hline
\end{tabular}
\end{table}
\end{landscape}

@@

Group-based hypothesis testing was a recurring theme across topics, with
text descriptions varying based on method(s) used. One-way analysis of
variance also featured strongly in topic 4 (n = 10212; 8.8\%), combined
with common methods for performing post-hoc multiple comparisons.
Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
mentions of descriptive statistics for summarising continuous variables,
for example: ``Data are expressed as means ± standard error of the mean
(SEM).'' Sections assigned to this topic appeared to be expanded
versions of topics 1 and 9. Examples of boilerplate text were in the
form of descriptive statistics followed by hypothesis testing, for
example, Student's t-test, Mann Whitney U or one-way analysis of
variance.

Examples of boilerplate text for selected topics are presented in Table
\ref{tab:plos-example-boilerplate}. Detailed findings based on cosine
similarity scores are provided in Supplementary File 2.

@@

\caption{\label{fig:anzctr-missing}ANZCTR search results.}\label{fig:unnamed-chunk-7}
\end{figure}

@@

{\centering \includegraphics[width=0.7\linewidth]{figures/anzctr.wordclouds}

@@

\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-8}
\end{figure}

@@

>>>>

Two statisticians on this paper (AB and NW) have heard researchers admit
that they have copied-and-pasted their statistical methods sections from
other papers, regardless of whether they are appropriate. The aim of
this paper is to use text-mining methods to analyse the content of
statistical methods sections included in common research outputs.
Results are then evaluated to estimate the extent that researchers are
using cut-and-paste or `boilerplate' statistical methods sections.
Boilerplate text is that ``which can be reused in new contexts or
applications without significant changes to the original''
\citep{Wikipedia}. Use of these methods sections indicates that little
thought has gone into the statistical analysis.

@@

We used two openly available data sources to find statistical methods
sections: research articles published in \emph{PLOS ONE} and study
protocols registered on the Australian and New Zealand Clinical Trials
Registry (ANZCTR). Data sources were chosen as examples of common
research outputs that should include descriptions of statistical methods that
were used, or are planned, for analysing study outcomes.

@@

\emph{PLOS ONE} is a large open access journal that publishes original
research across a wide range of scientific fields. Article submissions
are handled by an academic editor who selects peer reviewers based on
their self-nominated areas of expertise. Submissions do not undergo
formal statistical review. Instead, reviewers are required to assess
submissions against several publication criteria, including whether:
``Experiments, statistics, and other analyses are performed to a high
technical standard and are described in sufficient detail''
\citep{PLOS}. All reviewers are asked the question: ``Has the
statistical analysis been performed appropriately and rigorously?'',
with the possible responses of ``Yes'', ``No'' and ``I don't know''.

@@

\emph{Step 1: Targeted API searches}. API searches were completed using
the `rplos' package \citep{rplos}. Search queries targeted analysis-related terms, combining 
the words ``data'' or ``statistical'' with one of:
``analysis'', ``analyses'', ``method'', ``methodology'' or
``model(l)ing''. We allowed terms to appear anywhere in the article, to allow for the possibility of relevant text being placed in different
sections, for example, in the \emph{Material and Methods} section versus
\emph{Results}. Search results were indexed by a unique Digital Object
Identifier (DOI). Attribute data collected per DOI included journal
volume and subject classification(s).

\emph{Step 2: Partial matching on section headings}. Full text XML data
were downloaded and combined into a single dataset, organised by DOI and
subsection heading(s). Since \emph{PLOS ONE} does not prescribe
standardised headings to preface statistical methods sections, we
performed partial matching on available headings against frequently used
terms in initial search results: `Statistical analysis', `Statistical
analyses', `Statistical method', `Statistics', `Data analysis' and `Data
analyses'.

@@

 Text cleaning aimed to standardise notation and
statistical terminology, whilst minimising changes to article style and
formatting. Datasets were cleaned using the `textclean package'
\citep{textclean}. \emph{R} code used for data extraction and cleaning is
available from \url{https://github.com/agbarnett/stats_section}.

@@

We compiled an extensive list of statistical terms to standardise
descriptions of statistical methods reported across both datasets. An
initial list was compiled by calculating individual word frequencies and
identifying relevant terms that appeared at least 100 times. Further
terms were sourced from index searches of three statistics textbooks
\citep{Diggle2013,Bland2015,Dobson2018}. The final list
is provided in Supplementary File 1. Plurals (e.g., `chi-squares')
unhyphenated (e.g., `chi square') and combined (e.g.~`chisquare') terms
were transformed to singular, hyphenated form (e.g., `chi-square').
Common statistical tests were also hyphenated (e.g., `hosmer lemeshow'
to `hosmer-lemeshow').

@@

Text clustering is known to face the curse of dimensionality due to the
high number of terms in doc\(\times\)term matrix representation
\citep{sutanto2018fine,aggarwal2012mining}. Therefore
text-based methods based on distance, density or probability face
difficulties
\citep{mohotti2018corpus,mohotti2019concept,park2018examining}.
Specifically, the distance difference between near and far points
becomes negligible in high-dimensional data \citep{aggarwal2012mining}.
This directly affects the distance-based methods such as
\textit{k}-means \citep{jain} in accurately identifying the common
subgroups. In addition, the sparseness of this high dimensional matrix
representation does not allow to differentiate the user groups based on
density differences
\citep{mohotti2018efficient,mohotti2018corpus}.

Non-negative Matrix factorization (NMF) which maps the high-dimensional
data to a lower-dimensional space has been found to provide an effective
solution by allowing to form clusters in the lower-dimensional space
\citep{luong2019clustering,kim2014algorithms}. NMF approximates
topic vectors in linear manner considering the context of terms.

@@

Within the \emph{PLOS ONE} dataset, statistical methods sections with
high topic coherence scores were used to investigate evidence of
boilerplate text, compared with other studies assigned to the same
topic. Comparisons were carried out by calculating Jaccard similarity
scores between studies at the sentence level. We chose the Jaccard score
as an intuitive measure, which summarises the similarity between two
sentences as the number of words common to each sentence (intersection),
divided by the number of words that appeared in either the strongest
match or the sentence being compared. Boilerplate text was defined based
on word count (plus or minus 3 words) and a Jaccard score of 0.9 or
higher. Results were transformed to lower case for the clustering, but examples
are given using the original capitalisation.

@@

{\centering \includegraphics[width=0.8\linewidth]{figures/plossummary.jpg}

@@

\caption{\label{fig:plos-n} A: Search results by PLOS ONE volume; B: word count per statistical methods section included in analysis (n = 111,73); C: subject classifications assigned to full-text records included in analysis}\label{fig:unnamed-chunk-3}
\end{figure}

@@

{\centering \includegraphics[width=0.8\linewidth]{figures/ploswordclouds.jpg}

@@

\begin{landscape}
\begin{table}

\caption{\label{tab:plos-example-boilerplate}Example boilerplate text from PLOS ONE dataset. Sentences marked with an asterisk $^{\dagger}$ correspond to the statistical methods section with the strongest match based on topic coherence score.}
\centering
\resizebox{1.15\textwidth}{!}{\begin{tabular}[t]{p{0.05\linewidth} p{0.80\linewidth} p{0.15\linewidth}}
\hline
Topic & Statistical methods text & Matching DOIs\\
\hline
1  & $^{\dagger}$Students t-test was used for statistical analysis & 29\\
 & $^{\dagger}$A p value of less-than 0.05 was considered statistically significant. & 1883\\
\cline{2-3}
 & Statistical analysis was performed using the students t-test & 380\\
  & p values of less-than 0.05 were considered significant. & 1526\\
\hline
3 & $^{\dagger}$GraphPad Prism (Graphpad Software, San Diego, CA) was used for all analyses. & 1941\\
\cline{2-3}
 & All statistical analysis was performed using Graphpad Prism software. & 1312\\
\hline
4 & $^{\dagger}$Significant differences were determined using analysis of variance (ANOVA) followed by Tukey post-hoc tests for multiple comparisons. & 2216\\
\cline{2-3}
 & Data are expressed as the mean ± SEM & 860\\
 & Statistical analysis was performed using one-way analysis of variance (ANOVA) test, followed by Dunnett’s multiple comparison post hoc test & 4516\\
 & P values less than 0.05 were considered significant & 2414\\
\hline
5 & $^{\dagger}$SPSS software version 17.0 (SPSS, Chicago, IL, USA) was used for statistical analysis. & 441\\
\cline{2-3}
 & The results are presented as the mean ± SEM. & 420\\
 & Data were analyzed by one-way ANOVA and LSD tests using the SPSS 13.0 software (SPSS Inc., Chicago, IL, USA). & 19\\
 & The difference was considered statistically significant at P$<$0.05. & 2555\\
\hline
6 & $^{\dagger}$All results are expressed as means ± standard deviation (± SD). & 102\\
\cline{2-3}
 & Data are expressed as mean ± Standard error of the mean (SEM) & 604\\
 & Statistical analysis of the data was performed by the Student’s t-test & 444\\
 & A value of P$<$0.05 was considered statistically significant. & 1786\\
\hline
9 & $^{\dagger}$Statistical significance was determined by Students t-tests. & 49\\
 & $^{\dagger}$$^{*} P<0.05$, $^{**} P = <0.01$, $^{***} P<0.001$. & 6\\
\cline{2-3}
 & Data are presented as the mean ± SD. & 378\\
 & Statistical analysis was performed using the Students unpaired t-test. & 752\\
 & Differences were considered statistically significant at $^{*}p<0.05$, $^{**}p<0.01$ and $^{***}p<0.001$. & 413\\

\hline
\end{tabular}}
\end{table}
\end{landscape}

@@

Group-based hypothesis testing was a recurring theme across topics, with
text varying based on method(s) used. One-way analysis of
variance featured strongly in topic 4 (n = 10212; 8.8\%), combined
with common methods for performing post-hoc multiple comparisons, for example, Tukey's test and Dunnett's test.
Frequently occurring words in topic 6 (n = 4764; 4.1\%) reflected
mentions of descriptive statistics for summarising continuous variables,
for example: ``Data are expressed as means ± standard error of the mean
(SEM).'' Sections assigned to this topic appeared to be expanded
versions of topics 1 and 9. Examples of boilerplate text were in the
form of descriptive statistics followed by hypothesis testing, for
example, Student's t-test, Mann Whitney U or one-way analysis of
variance.

Examples of boilerplate text for selected topics are presented in Table
\ref{tab:plos-example-boilerplate}.

@@

\caption{\label{fig:anzctr-missing}ANZCTR search results.}\label{fig:unnamed-chunk-8}
\end{figure}

@@

{\centering \includegraphics[width=0.7\linewidth]{figures/anzctrwordclouds}

@@

\caption{ANZCTR topic clouds.}\label{fig:unnamed-chunk-9}
\end{figure}

@@

Targeted searches for the term `statistician' revealed
instances across all topics where no methods were specified. Instead,
all analyses were the responsibility of an appointed statistician, for
example, ``A statistician employed by hospital was used'' and ``Pilot
study at this point will use a statistician professionally to determine
sample size calculations as required''.
